Convex Surrogates via Polyhedral Losses

### 2019-02-02 cvx-fin.tex
started abstract

\usetikzlibrary{calc}
\newcommand{\Comments}{1}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
  \todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\ifnum\Comments=1               % fix margins for todonotes
  \setlength{\marginparwidth}{1in}
\fi

@@

\begin{abstract}
  \bo{TODO - iterate!}
  Convex surrogates are sweet.
  Given a loss for a classification-like problem, there are two natural approaches to design convex surrogates.
  First, one may attempt to map each prediction to a low-dimensional vector, and try to find a convex loss in that space with the right calibration.
  Second, one may simply try to find a surrogate within the class of piecewise-linear convex, or polyhedral, losses.
  We show an equivalence between these two approaches, and \raf{more stuff}.
  We show that every loss with a finite number of predictions has a convex surrogate in the above sense using one fewer dimension that the number of outcomes, and give a full characterization of the losses needing only $d$ dimensions for such a surrogate.
  We then apply this characterization to show novel lower bounds for abstain loss, demonstrating the power of our techniques over \raf{check this} alternatives such as feasible subspace dimension.
\end{abstract}
\begin{keywords}%
  information elicitation, proper scoring rules, surrogate loss functions
\end{keywords}

@@

WLOG, suppose $v_1 = (0,1)$ and they are numbered clockwise with $\|v_1\| \geq \|v_j\| (\forall j \neq i)$.
Let $\theta_{i,j}$ be the clockwise angle between $v_i$ and $v_j$.
The vectors point to the sides of an irregular pentagon $T^* = \{w : Vw \leq \vec{1}\}$.
Let $w_{1,2}$ be the corner of this pentagon between $v_1$ and $v_2$, and so on.
In our case, each $T_y$ is the intersection of $T^*$ with a halfspace, $T_y = \{w \in T^* : \inprod{v_y}{w} \leq -1\}$.
So in particular, $(V,B_1)$ is tight only if all three vertices $w_{23},w_{34},w_{45}$ satisfy $\inprod{v_1}{w} \leq -1$.
(If for example $\inprod{v_1}{w_{23}} > -1$, then the constraint $\inprod{v_2}{w} \leq 1$ is not tight, i.e. there is a separation between it and $T_1$.)
Because $v_1$ is the largest vector by assumption, this implies $\theta_{13} \geq 180^{\circ}$, but it also implies $\theta_{41} \geq 180^{\circ}$, which is a contradiction (both vectors would have to lie directly opposite $v_1$, but then they would be equal).

\begin{center}
\begin{tikzpicture}[scale=2.0]
 \path[fill=gray, opacity=0.2] (0.95105651629,0.30901699437) coordinate(p1) --  ++(180:1) coordinate(p5)
 -- ++(252:1) coordinate(p4) --
 ++(324:1) coordinate(p3) --  ++(396:1) coordinate(p2);

 \coordinate (ctr) at (barycentric cs:p1=1,p2=1,p3=1,p4=1,p5=1);
 \foreach \X [count=\Y] in {2,...,6}
 {\ifnum\X=6
   \path (p\Y) -- (p1) coordinate[pos=0.0](a\Y) coordinate[pos=1.0](a1)
   coordinate[pos=0.5](m1);
   \draw (a\Y) -- (a1);
   \draw[-latex] (ctr) -- ($ (m1)!0.0cm!90:(p1) $) node[pos=1.2]{$v_{\Y}$};
  \else
   \path (p\Y) -- (p\X) coordinate[pos=0.0](a\Y) coordinate[pos=1.0](a\X)
   coordinate[pos=0.5](m\X);
   \draw (a\Y) -- (a\X);
   \draw[-latex] (ctr) -- ($ (m\X)!0.0cm!90:(p\X) $) node[pos=1.2]{$v_{\Y}$};
  \fi}
  \draw[fill=blue, opacity = 0.2] (-1.0,-0.95105) -- (-1.0,-1.52) -- (2.0,-1.52) -- (2.0,-0.95105) -- cycle;
  \node[rotate=54, xshift=0.5cm] at (ctr) {\tiny$\theta_{5,1}$};
  \draw (0.555, -0.35) arc (18:90:0.12);
  \node at (0.5,-1.32) {\tiny$T_5 := \{w : \langle v_5, w\rangle \leq -1 \}$};
  \draw[fill=blue, opacity = 0.8] (m3) -- (m4) -- (a3) -- cycle;
\end{tikzpicture}
\end{center}
\end{proof}

>>>>

\newcommand{\Comments}{1}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
  \todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\ifnum\Comments=1               % fix margins for todonotes
  \setlength{\marginparwidth}{1in}
\fi

@@

\begin{abstract}
  \bo{TODO - iterate!}
  We formalize and study two approaches to designing convex surrogate loss functions for problems such as classification or ranking.
  First, one can attempt to ``embed'' each prediction (label, ranking, etc) as a point in $\reals^d$ and design a convex loss function on that space.
  Second, one can use a piecewise-linear convex, or polyhedral loss.
  We prove an equivalence: if embedding is possible, than it can be done with a polyhedral loss, and 
%  Convex surrogates are sweet.
%  Given a loss for a classification-like problem, there are two natural approaches to design convex surrogates.
%  First, one may attempt to map each prediction to a low-dimensional vector, and try to find a convex loss in that space with the right calibration.
%  Second, one may simply try to find a surrogate within the class of piecewise-linear convex, or polyhedral, losses.
%  We show an equivalence between these two approaches, and \raf{more stuff}.
%  We show that every loss with a finite number of predictions has a convex surrogate in the above sense using one fewer dimension that the number of outcomes, and give a full characterization of the losses needing only $d$ dimensions for such a surrogate.
%  We then apply this characterization to show novel lower bounds for abstain loss, demonstrating the power of our techniques over \raf{check this} alternatives such as feasible subspace dimension.
\end{abstract}
\begin{keywords}%
  information elicitation, proper scoring rules, surrogate loss functions
\end{keywords}

@@

WLOG, suppose $v_1 = (0,1)$ and they are numbered clockwise with $\|v_1\| \geq \|v_j\| (\forall j \neq i)$.
Let $\theta_{i,j}$ be the clockwise angle between $v_i$ and $v_j$.
The vectors point to the sides of an irregular pentagon $T^* = \{w : Vw \leq \vec{1}\}$.
Let $w_{1,2}$ be the corner of this pentagon between $v_1$ and $v_2$, and so on.
In our case, each $T_y$ is the intersection of $T^*$ with a halfspace, $T_y = \{w \in T^* : \inprod{v_y}{w} \leq -1\}$.
So in particular, $(V,B_1)$ is tight only if all three vertices $w_{23},w_{34},w_{45}$ satisfy $\inprod{v_1}{w} \leq -1$.
(If for example $\inprod{v_1}{w_{23}} > -1$, then the constraint $\inprod{v_2}{w} \leq 1$ is not tight, i.e. there is a separation between it and $T_1$.)
Because $v_1$ is the largest vector by assumption, this implies $\theta_{13} \geq 180^{\circ}$, but it also implies $\theta_{41} \geq 180^{\circ}$, which is a contradiction (both vectors would have to lie directly opposite $v_1$, but then they would be equal).
\end{proof}
